Optimal strategies in mental rotation tasks

### 2013-05-13 man/final-project.tex
Results writeup for everything except bayesian quadrature model

>>>>

We evaluated each model's performance on 50 randomly generated shapes
(such as $X_a$ in Figure \ref{fig:stimuli}). For each shape, we picked
a random angle $R$ and generated two stimuli: a $h_0$ stimulus, by
reflecting $X_a$ across the $y$-axis and rotating it by $R$; and a
$h_1$ stimulus, by simply rotating $X_a$ by $R$.

We considered three metrics of performance in particular:
\begin{itemize}
\item \textit{Rotations}: for stimuli where $h_1$ is correct, how
  correlated are the true angles of rotation with the rotations
  performed by the model? This is defined to be the Pearson's
  correlation coefficient $\rho$ between $R$ and $\vert
  \mathbf{R}\vert$. Figure \ref{fig:rotations} shows individual points
  corresponding to true rotation vs. rotation by the model, for each
  model.
\item \textit{Responses}: how accurate is the model at choosing the
  correct hypothesis? This is defined to be the mean error
  ($\mathrm{ME}$), or fraction of times the model picks the incorrect
  hypothesis.
\item \textit{Estimates of $Z$}: how accurate is the model's estimate
  of $Z$? This is defined to be the mean squared error
  ($\mathrm{MSE}$) between the model's estimate of $Z$ and the ``gold
  standard'' value, where the error has scale such that
  $\mathrm{MSE}=0$ indicates no error, and $\mathrm{MSE}=1$ indicates
  maximum error. Figure \ref{fig:accuracy} shows plots of the true
  (gold standard) value of $Z$ vs. the model's estimate, for each
  model.
\end{itemize}

\paragraph{Gold standard}

The gold standard always evaluates $S$ at all angles, so the
correlation between it and the true angle of rotation is
undefined. The mean error of responses was $\mathrm{ME}=0.01$,
indicating that $Z$ is a sufficient approximation\footnote{The one
  error comes from a $h_0$ stimulus which is highly
  symmetric. Although it is distinct enough for the author to
  correctly judge when both images are viewed side by side, it remains
  to be seen whether human participants, viewing the images
  sequentially would also make the correct judgment.}  of $\int_R
p(X_b\ \vert\ X_R)p(R)\ \mathrm{d}R$.

\paragraph{Na\"ive model}

The correlation between the number of rotations computed by the
na\"ive model and the true angle of rotation was $\rho=-0.11$. The
shape of the data is more interesting (see Figure \ref{fig:rotations},
right panel): the na\"ive model actually corresponds quite well to the
true angle of rotation for $R<\frac{\pi}{2}$ ($\rho=0.63$). This is
unsurprising, because the closer the true angle is to zero, the less
the model has to rotate, and the less likely it will get stuck on
local maxima. Thus, it is more likely to locat the global maximum,
which generally corresponds to the true angle of rotation. For
$R>\frac{\pi}{2}$, we see an increasing tendency to under-rotate
($\rho=-0.42$); this is likely because it finds a local maxima and
ends prematurely.

The na\"ive model's response error rate was better than chance, but
still quite high, with $\mathrm{ME}=0.23$. Closer inspection reveals
that the bulk of this comes from $h_1$ stimuli ($\mathrm{ME}=0.36$
vs. $\mathrm{ME}=0.1$ for $h_0$ stimuli). This is, as above, probably
related to finding only local maxima: if the model finds a local
maxima which is low enough, the area under the estimated curve will
not be large enough to accept $h_1$.

This is reflected in the accuracy in estimating $Z$ as well
($\mathrm{MSE}=0.14$). Figure \ref{fig:accuracy} (leftmost panel)
shows individual estimates for each stimulus, color-coded by true
hypothesis. The model underestimates $Z$ for significant portion of
the $h_1$ stimuli. On the other $h_1$ stimuli, the na\"ive model
actually tends to overestimate because linear interpolation does not
account for symmetry of global maxima (\TODO{figure?}).

\paragraph{Parametric (Von Mises) model}

Like the na\"ive model, the parametric model uses hill-climbing as a
sampling strategy, so they have identical correlations with the true
angle of rotation. However, given the same $\mathbf{R}$, the
parametric model estimates $Z$ differently, thus giving rise to
different response and estimated $Z$ error rates. The response error
for the parametric model was higher than the na\"ive model, with
$\mathrm{ME}=0.26$. Breaking this down, we see that the error for
$h_1$ stimuli is $\mathrm{ME}=0.5$: in other words, the parametric
model is at chance when determining whether two identical shapes are
the same. This is because it never overestimates (see Figure
\ref{fig:accuracy}, center panel). So, it accurately estimates $Z$ for
stimuli which the na\"ive model would overestimate
($\mathrm{MSE}=0.07$), but still has poor performance when $Z$ is
underestimated. Indeed, if we exclude stimuli for which the nai\"ive
model overestimates ($Z\geq 3$), the error for the na\"ive model
lowers to match that of the parametric model.

\paragraph{Nonparametric (Bayesian Quadrature) model}
